### 2022-04-23 conclusions.tex
Refactor the latex code

%! Author = jincheng
%! Date = 2022/4/23

\section{Conclusions}
\label{sec:conclusions}
In this paper, we propose the idea that using target speaker's pitch as an auxiliary feature to improve the performance of target speaker separation. A target pitch extraction model is built and the target pitch information is incorporated with the TSS models in both simply concatenation and joint training strategies. We found that the target pitch information could improve the separation performance even though the pitch precision is not high enough yet. While the performance of concatenation with ground-truth pitch information show the great potency of this approach. The joint training approach yields better performance than simple concatenation. We also explore if joint training would bring improvement to the target pitch extraction, the result shows no obvious help. In the future work, we will continue to improve the precision of target pitch extraction and do more experiments on large scale models to validate the proposed methods.