### 2022-04-23 acknowledgements.tex
Add acknowledgements part

%! Author = jincheng
%! Date = 2022/4/23

\section{ACKNOWLEDGEMENTS}\label{sec:acknowledgements}
The first I really want to thank is Prof. Ming Li. I applied to joint Prof. Ming Li's Lab during the summer of my freshman year.
Later on, under the guidance and resources provided by the professor, I started my research in the field of speech analysis technology and speech separation.