### 2022-04-23 acknowledgements.tex
Update the acknowledgements part

%! Author = jincheng
%! Date = 2022/4/23

\section{ACKNOWLEDGEMENTS}\label{sec:acknowledgements}
First and foremost, I would like to thank my research mentor Professor Ming Li who provided huge help to me in the past three years.
I first met the professor was in the class of CS 101.
Later on I applied to join Prof. Ming Li's Speech and Multimodal Intelligent Information Processing (SMIIP) Lab during the summer of my freshman year.
From that time, under the guidance and resources provided by the professor, I started my research in the field of speech analysis technology and speech separation.