### 2022-04-24 background.tex
Update metrics SISNR

%! Author = jincheng
%! Date = 2022/4/24

\section{Background}\label{sec:background}

In this section, we give a detailed background about what is target speaker separation (TSS) task, what kinds of metrics can be used to define the performance, what kinds of techniques existed can be used to do TSS task.

\subsection{What is TSS task}\label{subsec:what-is-tss-task}

Assume there are $N$ speakers talking simultaneously, they are called $s_1, s_2, \dots s_N$. The speech utterance for each speaker is $U_i$ while $i\in\left[ 1, N \right]$. Assume the noise is $e$.
So the mixed utterance $U$ is as equation~\ref{eq:mixed}.
\begin{equation}
    U = \sum\limits_{i=1}^N U_i + e\label{eq:mixed}
\end{equation}
For the mission of Target Speaker Separation (TSS) Task. Assume we only need to extract the voice of speaker $s_n$ and we have the reference utterance from $s_n$, denoted as $R_n$.
The task is given $U = \sum\limits_{i=1}^N U_i + e$ and $R_n$, we need to get $U_n$.

\subsection{Metrics}\label{subsec:metrics}
\begin{itemize}
    \item scale-invariant source-to-noise ratio (SI-SNR)

    SI-SNR is proposed in~\cite{tasnet}.
    \begin{equation}
        \begin{aligned}
            s_{\text{target}} &= \frac{\left< \hat{s}, s \right>s}{\| s \|^2} \\
            s_{\text{noise}} &= \hat{s} - s_{\text{target}} \\
            \text{SI-SNR} &\coloneqq 10\log_{10}\frac{\| s_{\text{target}} \|^2}{\| e_{\text{noise}} \|^2}
        \end{aligned}\label{eq:si-snr}
    \end{equation}
\end{itemize}

\subsection{Common Methods}\label{subsec:common-methods}