### 2022-04-24 background.tex
Update background formula

%! Author = jincheng
%! Date = 2022/4/24

\section{Background}\label{sec:background}

In this section, we give a detailed background about what is target speaker separation (TSS) task, what kinds of metrics can be used to define the performance, what kinds of techniques existed can be used to do TSS task.

\subsection{What is TSS task}\label{subsec:what-is-tss-task}

Assume there are $N$ speakers talking simultaneously, they are called $s_1, s_2, \dots s_N$. The waveform of speech utterance for each speaker is $U_i\left( t \right)$ while $i\in\left[ 1, N \right]$. Assume the noise is $e\left( t \right)$.
So the mixed utterance $X\left( t \right)$ is as equation~\ref{eq:mixed}.
\begin{equation}
    X\left( t \right) = \sum\limits_{i=1}^N U_i\left( t \right) + e\left( t \right)
    \label{eq:mixed}
\end{equation}
For the mission of Target Speaker Separation (TSS) Task. Assume we only need to extract the voice of speaker $s_n$ from $X\left( t \right)$ and we have the reference utterance from $s_n$, denoted as $R_n$.
The task is given $X\left( t \right) = \sum\limits_{i=1}^N U_i\left( t \right) + e\left( t \right)$ and $R_n$, we need to get $U_n\left( t \right)$.

\subsection{Metrics}\label{subsec:metrics}

Using meaningful metrics to define the performance of the separation system is important.

\begin{itemize}
    \item source-to-distortion ratio (SDR)

    SDR is defined in~\cite{performance_measurement}, and is as equation~\ref{eq:sdr}.
    \begin{equation}
        \text{SDR} \coloneqq 10\log_{10}\frac{\| s_{\text{target}} \|^2}{\| e_{\text{interf}} + e_{\text{noise}} + e_{\text{artif}} \|^2}
        \label{eq:sdr}
    \end{equation}

    \item scale-invariant source-to-noise ratio (SI-SNR)

    SI-SNR is proposed in~\cite{tasnet}, and is as equation~\ref{eq:si-snr}.
    \begin{equation}
        \begin{aligned}
            s_{\text{target}} &= \frac{\left< \hat{s}, s \right>s}{\| s \|^2} \\
            e_{\text{noise}} &= \hat{s} - s_{\text{target}} \\
            \text{SI-SNR} &\coloneqq 10\log_{10}\frac{\| s_{\text{target}} \|^2}{\| e_{\text{noise}} \|^2}
        \end{aligned}\label{eq:si-snr}
    \end{equation}
\end{itemize}

\subsection{Methodology}\label{subsec:methodology}