### 2022-04-25 background.tex
Check grammar of background

%! Author = jincheng
%! Date = 2022/4/24

\section{Background}\label{sec:background}

In this section, we give a detailed background about what is target speaker separation (TSS) task, what kinds of metrics can be used to define the performance and what kinds of techniques existed can be used to do TSS task.

\subsection{What is TSS task}\label{subsec:what-is-tss-task}

Assume there are $N$ speakers talking simultaneously, they are called $s_1, s_2, \dots s_N$. The waveform of speech utterance for each speaker is $U_i\left( t \right)$ while $i\in\left[ 1, N \right]$. Assume the noise is $e\left( t \right)$.
So the mixed utterance $X\left( t \right)$ is as equation~\ref{eq:mixed}.
\begin{equation}
    X\left( t \right) = \sum\limits_{i=1}^N U_i\left( t \right) + e\left( t \right)
    \label{eq:mixed}
\end{equation}
One waveform example is as figure~\ref{fig:waveform}.
\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{img/waveform}
    \caption{One example of how the waveform looks like}
    \label{fig:waveform}
\end{figure}

For the mission of the Target Speaker Separation (TSS) Task. Assume we only need to extract the voice of speaker $s_n$ from $X\left( t \right)$ and we have the reference utterance from $s_n$, denoted as $R_n$.
The task is given $X\left( t \right) = \sum\limits_{i=1}^N U_i\left( t \right) + e\left( t \right)$ and $R_n$, we need to get $U_n\left( t \right)$.

\subsection{Metrics}\label{subsec:metrics}

Using meaningful metrics to define the performance of the separation system is important.
To evaluate the performance of the separation system, the metric should be designed to objectively represent ``how good'' the extracted waveform is.
The idea of projecting one high dimensional vector onto another high dimensional vector comes to mind.

\begin{itemize}
    \item source-to-distortion ratio (SDR)

    SDR is defined in~\cite{performance_measurement}, and is as equation~\ref{eq:sdr}.
    \begin{equation}
        \text{SDR} \coloneqq 10\log_{10}\frac{\| s_{\text{target}} \|^2}{\| e_{\text{interf}} + e_{\text{noise}} + e_{\text{artif}} \|^2}
        \label{eq:sdr}
    \end{equation}

    \item scale-invariant source-to-noise ratio (SI-SNR)

    SI-SNR is proposed in~\cite{tasnet}, and is as equation~\ref{eq:si-snr}.
    \begin{equation}
        \begin{aligned}
            s_{\text{target}} &= \frac{\left< \hat{s}, s \right>s}{\| s \|^2} \\
            e_{\text{noise}} &= \hat{s} - s_{\text{target}} \\
            \text{SI-SNR} &\coloneqq 10\log_{10}\frac{\| s_{\text{target}} \|^2}{\| e_{\text{noise}} \|^2}
        \end{aligned}\label{eq:si-snr}
    \end{equation}
\end{itemize}

\subsection{Methodology}\label{subsec:methodology}

To extract the target speaker's voice from the mixed utterance, we first need to extract the information of the target speaker.
One common way is to use a vector to represent the target speaker's information, and we call it the speaker embedding.
There are many ways and many different speaker embeddings, such as i-vector based on simple factor analysis in~\cite{i-vector}, a speaker embedding based on end-to-end neural network in~\cite{deep_speaker}, x-vector in~\cite{x-vector}, d-vector in~\cite{d-vector}.
As Transformer architecture~\cite{attention} being existed, there are some Transformer encoder based speaker embeddings happen, such as s-vector in~\cite{s-vector}.

After extracting the speaker embedding, there are basically two different ways to do the separation task, one is frequency-domain methods, and another is time-domain methods.

\subsubsection{Frequency-domain methods}

For the frequency-domain methods, the central part is processing the magnitude spectrogram. Here gives an example in figure~\ref{fig:spec_exp}.
\begin{figure}
    \centering
    \stackunder[5pt]{\includegraphics[width=0.9\linewidth]{img/spec_exp_mix}}{(a)}
    \stackunder[5pt]{\includegraphics[width=0.9\linewidth]{img/spec_exp_tar}}{(b)}
    \stackunder[5pt]{\includegraphics[width=0.9\linewidth]{img/spec_exp_est}}{(c)}
    \caption{Magnitude spectrogram example of 1650-167613-0006. (a) Magnitude spectrogram of the mixed utterance; (b) Magnitude spectrogram of the target speaker utterance; (c) Magnitude spectrogram of the estimated target speaker utterance. }
    \label{fig:spec_exp}
\end{figure}

For each utterance, by doing Fourier transform, it can be converted into magnitude spectrogram and phase spectrogram.
For the frequency-domain methods, the central part is to mask the magnitude spectrogram.
In (a), it is the magnitude spectrogram of the mixed utterance which contains two speakers, and (b) is the magnitude spectrogram of the target speaker's utterance.
The goal is by giving (a) and the target speaker's reference audio, we want to get (b). (c) is our system's estimated magnitude spectrogram of the target speaker's utterance, and it is very visually similar to (b).
For the frequency-domain methods, the mask is estimated, and is dot-multiplied to the mixed magnitude spectrogram for getting the estimated one.
The whole process can be described in figure~\ref{fig:sep_process}.
\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{img/sep_process}
    \caption{Frequency-domain methods general description}
    \label{fig:sep_process}
\end{figure}

There are many methods using this framework, such as deep clustering~\cite{deep_clustering}, which trains a deep network to produce spectrogram embeddings.
The estimated mask is binary, and the projection process is achieved by the deep neural network, and the objective function it trains is as equation~\ref{eq:deep_clustering}.
\begin{equation}
    C_Y\left( V \right) = \| VV^T - YY^T \|_F^2
    \label{eq:deep_clustering}
\end{equation}

Also, permutation invariant training (PIT) in~\cite{PIT} is widely used in blind speaker separation.
One big problem of speaker-independent speech separation is the permutation of the output of the network.
PIT solves this problem, which outputs all the permutations and using the smallest mean squared error (MSE) as the optimization goal.
PIT structure is as figure~\ref{fig:PIT_structure}.
To better solve the issue of tracking the speaking person, one utterance-level algorithm uPIT in~\cite{uPIT} is proposed.
\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{img/PIT_structure}
    \caption{Permutation invariant training structure~\cite{PIT}}
    \label{fig:PIT_structure}
\end{figure}

To better solve the disadvantages of deep clustering, a new deep embedding clustering based algorithm DEF-DL~\cite{DEF-DL} is proposed.
DEF-DL uses a supervised network to replace K-means clustering algorithm, and uses this feature as the input of uPIT network to separate speech.
The whole structure is as figure~\ref{fig:def_dl_structure}.
\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.6\textwidth]{img/DEF-DL_structure}
    \caption{DEF-DL structure~\cite{DEF-DL}}
    \label{fig:def_dl_structure}
\end{figure}
For the target speaker separation, one classic structure is speaker encoder, speech encoder, separator, decoder framework, as shown in figure~\ref{fig:enc_sep_doc_arc}.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.7\textwidth]{img/voicefilter_struct}
    \caption{VoiceFilter structure~\cite{voicefilter}}
    \label{fig:voicefilter_struct}
\end{figure}

One example is VoiceFilter~\cite{voicefilter}, the structure is shown in figure~\ref{fig:voicefilter_struct}.
The speaker encoder is trained separately to get the speaker embedding.
For each iteration of training, the noisy audio, clean audio, and reference audio are given.
Operating Short-time Fourier transform (STFT) to both noisy audio and clean audio to get the magnitude spectrogram and phase spectrogram.
The magnitude spectrogram of the noisy audio is fed into the separator, for VoiceFilter, it is Convolutional Neural Network (CNN), Long short-term memory (LSTM)~\cite{LSTM} and fully-connected layer.
After the separator predicts the mask, dot multiplication is applied to the input magnitude spectrogram to get the predicted masked magnitude spectrogram.
Last, applying inverse STFT (iSTFT) to the masked magnitude spectrogram and the original phase spectrogram to get the enhanced audio.
The training loss of the network is mean squared error (MSE) between the masked magnitude spectrogram and the clean spectrogram from the clean audio, which is also very common in the frequency-domain methods.

However, applying iSTFT on the masked magnitude spectrogram and the original phase spectrogram to get the enhanced audio may affect the performance since the phase spectrogram is not processed.
There are some methods that can reconstruct phase information, such is WA-MISI in~\cite{WA-MISI} and Deep CASA in~\cite{Deep_CASA}.

\subsubsection{Time-domain methods}

To better solve the problem of the phase information is not matched to the masked spectrogram, many time-domain methods existed.
For the time-domain methods, the input and output are not magnitude spectrogram anymore.
These methods directly process the waveform and output the waveform.
Since the waveform contains both magnitude and phase information, the issue of mismatching phase can be solved.
There are some outstanding works of time-domain methods, such as Conv-TasNet~\cite{conv_tasnet}, which has three main modules: encoder, separator and decoder.
For the encoder, it uses 1-dimensional convolutional layer to replace STFT to encode the waveform, which forces the network to learn the parameters.
For the separator, it was fed the feature from the encoder and output the mask which is very similar to the frequency-domain methods.
Dot multiplication is applied to the mask and the output feature from the encoder.
Last but not least, the decoder which is a transpose 1-dimensional convolutional layer which takes the output from the separator to the audio signal.
Conv-TasNet directly uses Si-SNR~\ref{eq:si-snr} as the training loss, which can improve the performance and solve the issue of the mismatching phase information.
However Conv-TasNet can only focus on the segmented length of the audio, it cannot focus on the whole utterance.
A new system which is also a time-domain method called Dual-Path RNN (DPRNN) is proposed in~\cite{dprnn}, which can focus on the whole utterance and thus improve the performance.